---
title: "Mediation Analysis with Mplus via MplusAutomation"
output: html_notebook
---

## Setups
Load require packages. 
```{r message = F}
library(tidyverse)
library(MplusAutomation)
library(lavaan)
```

## ex2.1
Set up the Mplus specifications with MplusAutomation. Note that this is not a recommended way to conduct a mediation analysis, because the SE and CI are estimated by normal approximation (Delta method).
```{r}
sink(file = "ex2.1.inp")
cat('
data:     file = protest.txt;
variable: names = sexism liking respappr protest;
          usev = liking respappr protest;
analysis: estimator = mlr;
model:    liking on respappr protest;
          respappr on protest;
Model indirect:
          liking ind protest;
output:   cinterval;
  ', fill=T)
  sink()
```

Run the model.
```{r}
runModels("ex2.1.inp")
```

```{r}
ex2.1.fit <- readModels("ex2.1.out")
```

```{r}
ex2.1.fit$parameters$unstandardized
```

```{r}
ex2.1.fit$indirect
```

## ex2.1b
This is the same model as ex2.1, but estimates bootstrap CIs. Set up the Mplus specifications with MplusAutomation. Then, run the model with MplusAutomation.
```{r}
sink(file = "ex2.1b.inp")
cat('
data:     file = protest.txt;
variable: names = sexism liking respappr protest;
          usev = liking respappr protest;
analysis: estimator = ml; 
          bootstrap = 10000;
model:    liking on respappr protest;
          respappr on protest;
Model indirect:
          liking ind protest;
output:   cinterval(bootstrap);
', fill=T)
sink()
```

Run the model.
```{r}
runModels("ex2.1b.inp")
```

```{r}
ex2.1b.fit <- readModels("ex2.1b.out")
```

```{r}
ex2.1b.fit$parameters$unstandardized
```

```{r}
ex2.1b.fit$indirect
```
### ex2.1c
This is the same model as ex2.1b, and estimates bootstrap CIs, but manually create the mediation effect.
```{r}
sink(file = "ex2.1c.inp")
cat('
data:     file = protest.txt;
variable: names = sexism liking respappr protest;
          usev = liking respappr protest;
analysis: estimator = ml;
          bootstrap = 10000;
model:    liking on respappr (b1)
                    protest;
          respappr on protest (g1)
Model constraint: 
          new(prolik);
          prolik = g1 * b1;
output:   cinterval(bootstrap);
', fill=T)
sink()
```

Run the model.
```{r}
runModels("ex2.1c.inp")
```

```{r}
ex2.1c.fit <- readModels("ex2.1c.out")
```

```{r}
ex2.1c.fit$parameters$unstandardized
```


## Running by lavaan
Read Data. Data file `protest.txt` is read in. 
```{r}
protest.dat <- read.table("protest.txt")
colnames(protest.dat) <- c("sexism", "liking", "respappr", "protest")
```

Build a model.
```{r}
MA.mod01 <- '
liking ~ b1*respappr + protest
respappr ~ g1*protest
prolik := b1*g1
'
```

Run the model with bootstrap.
```{r}
MA.fit01 <- sem(MA.mod01, data=protest.dat, se="boot", bootstrap=10000)
```

Results
```{r}
summary(MA.fit01)
```

```{r}
parameterEstimates(MA.fit01, zstat=F, pvalue=F)
```


