---
title: "Class 2 - Mediation Analysis"
date: "`r Sys.Date()`"
output: 
  html_notebook:
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: true
---

```{r message=F, echo = FALSE, warning = FALSE, results='hide'}
# detach all libraries

detachAllPackages <- function() {
  basic.packages <- c("package:stats", "package:graphics", "package:grDevices", "package:utils", "package:datasets", "package:methods", "package:base")
  package.list <- search()[ifelse(unlist(gregexpr("package:", search()))==1, TRUE, FALSE)]
  package.list <- setdiff(package.list, basic.packages)
  if (length(package.list)>0)  for (package in package.list) detach(package,  character.only=TRUE)
}
detachAllPackages()

# function to load libraries

pkgTest <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[,  "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg,  dependencies = TRUE)
  sapply(pkg,  require,  character.only = TRUE)
}

# loading packages 


lapply(c("gsubfn"),pkgTest)
lapply(c("tidyverse"),pkgTest)
lapply(c("sjmisc"),pkgTest)
lapply(c("DT"),pkgTest)
lapply(c("lavaan"),pkgTest)
lapply(c("MplusAutomation"),pkgTest)
lapply(c("psych"),pkgTest)
library(gsubfn)
library(tidyverse)
library(sjmisc)
library(DT)
library(lavaan)
library(MplusAutomation)
library(psych)
```

### Mediation Analysis 

- The difference $c-c^\prime$ is considered as the mediation effect, and $c-c^\prime = a \times b$ or $c = a \times b - c^\prime$
- However, the equity holds only for a special case of mediation analysis 
  * Continuous outcome and linear regression 
- Therefore, it is not recommended to evaluate $c - c^\prime$

**Basic Mediation Analysis**

- Data from the discrimination study (Garcia et al., 2010) in Muthen et al. (2016; p.72), as well as Hayes (2013) 
- Subjects (129 females) were presented a story of hoow female attorney lost a promotion to a less qualified male due to discriminatory action 
- Subjects were randomly assigned to two different scenarios: 
  * (1) the female attorney takes no action, and (2) the female attorney protests 
- Variables 
  * Intervention variable (0 = no protest, 1 = protest) (`protest`) predictor 
  * How well the subject likes the female attorney outcome - 7-point likert (`liking`)
  * Perceived appropriateness of the response of the female attorney (`respappr`) mediator 
  * Beliefs in the pervasiveness of sex discrimination in society (`sexism`)
  
```{r}
protest_df <- read.table("protest.txt", sep="")
colnames(protest_df) <- c("sexism", "liking", "respappr", "protest")
psych::describe(protest_df)
```

- By substituting the second equation below to the first equation below, we can get the combined equation:  

Equation 1:
$$ y_i = \beta_0 + \beta_1m_i + \beta_2x_i+\epsilon_{yi}$$
Equation 2: 
$$ m_i = \gamma_0 + \gamma_1x_i+\epsilon_{mi}$$
Combined equation:
$$ y_i = \beta_0 + \beta_1(\gamma_0 + \gamma_1x_i + \epsilon_{mi})+\beta_2x_i+\epsilon_{yi}$$

- Example 1
  * Note that by default, Mplus utilizes a method to estimate a standard error of nonlinearly transformed parameters called the "Delta method". 
  * Delta method analytically approximates the target standard error by assuming the normality of the sampling distribution of the parameter(s). 
    - The indirect effect is a multiplication of two parameters 
  * Mplus syntax for the `model` section: `DV on IV1 IV2` and so on for however many IVs there are 
  
```{r}
sink("ex2.1.inp")
cat('
data:     file = protest.txt; 

variable: names = sexism liking respappr protest; 
          usev = liking respappr protest; 
          
analysis: estimator = mlr; 

model:    liking on respappr protest; 
          respappr on protest; 
          
Model indirect: 
          liking ind protest; 
          
output:   cinterval;
')
sink()

runModels("ex2.1.inp")

modelfit1 <- readModels("ex2.1.out")

modelfit1$parameters$unstandardized
```

**Basic Mediation Analysis with Bootstrap**

```{r}
sink("ex2.1b.inp")
cat('
data:     file = protest.txt; 

variable: names = sexism liking respappr protest; 
          usev = liking respappr protest; 
          
analysis: estimator = ml; 
          bootstrap = 10000;

model:    liking on respappr protest; 
          respappr on protest; 
          
Model indirect: 
          liking ind protest; 
          
output:   cinterval(bootstrap);
')
sink()

runModels("ex2.1b.inp")

modelfit2 <- readModels("ex2.1b.out")

modelfit2$parameters$unstandardized
```

**Another way to set up mediation analysis** 

```{r}
sink("ex2.1c.inp")
cat('
data:     file = protest.txt; 

variable: names = sexism liking respappr protest; 
          usev = liking respappr protest; 
          
analysis: estimator = ml; 
          bootstrap = 10000;

model:    liking on respappr (b1)
                     protest; 
          respappr on protest (g1); 
          
Model constraint: 
          new(prolik); 
          prolik = g1 * b1;
          
output:   cinterval(bootstrap);
')
sink()

runModels("ex2.1c.inp")

modelfit3 <- readModels("ex2.1c.out")

modelfit3$parameters$unstandardized
```

**Basic Mediation Analysis with** `lavaan`

```{r}
MA.mod01 <- '
liking ~ b1*respappr + protest 
respappr ~ g1*protest 
prolik := b1*g1
'
```

Run the model with bootstrap. 
```{r}
MA.fit01 <- sem(MA.mod01, data=protest_df, se="boot", bootstrap=10000)
```

```{r}
parameterEstimates(MA.fit01, zstat=F, pvalue=F)
```

**Mediation Analysis with Covariates** 

- Data from the Maternal Health Project (Day et al., 1994) is from a sample of mothers who drank at least three drinks a week duringn the first trimester of their pregnancy + mothers who drank alcohol less often ($n$ = 594). 
- Variables
  * Alcohol use (`momcig3`)
  * Cigarette use (`momalc3`)
  * Child's gender (`male`)
  * Ethnicity (`black`)
  * Head circumference at birth (`hcirc0`)
  * Head circumference at 36 monoths (`hcirc36`)

Equation 1: 
$$
y_i = \beta_0+\beta_1 m_i+\beta_2z_{1i}+\beta_3z_{2i}+\epsilon_{yi}
$$
$$
m_i=\gamma_0+\gamma_1x_{1i}+\gamma_2x_{2i}+\gamma_3z_{1i}+\gamma_4z_{2i}+\epsilon_{mi}
$$

**Mediation Analysis with multiple Mediators**

Both direct effect annd mediation effect depends on the value of $z_i$ as a covariate (i.e., moderated by the covariate)

If the moderation value is lower than .15, mediation effect is not significant as it's within the 95% band around 0. Within .15~+.85 would indicate significance, outside would be not significant. 


**One factor FA Model**

$$x_{ij}=\mu_i+\lambda_{i}\xi_{j}+\delta_{ij}$$
- This model is a set of more than one regression equation 
  * We have aas many regression equations as the number of items 
  * For example, let's say we have 3 items. Then, 
    - $x_{1j} = \mu_1 + \lambda_1\xi_j+\delta_{1j}$
    - $x_{2j} = \mu_2 + \lambda_2\xi_j+\delta_{2j}$
    - $x_{3j} = \mu_3 + \lambda_3\xi_j+\delta_{3j}$
    
- We can graphically represent this model (a set of three regression equations) as follows. 
  * It is typical to indicate observed variables by rectangles 
- The predictor $\xi$ is a latent factor and does not have any fixed scale 
- Traditionally, it has been typical to scale the means of $\xi$ and $x$ to be 0. 

**CFA model** 

- Model specification is needed for CFA 
  * A specific model needs to be specified, including: 
     - The number of observed indicators and latent factors 

$$ x_{1j} = \lambda_{11}\xi_1+0\xi_2+\delta_{1j}$$
$$ x_{2j} = \lambda_{21}\xi_1+0\xi_2+\delta_{2j}$$
- For ESEM, the 0 constraints would be removed 
  * Factor loadings for $\xi$ would be small but included 
- Requirements for identification of CFA model 
  * Just like a path analysis model, the number of parameters in the model has to be the same or fewere than the unique variances and covariances ini the data covariance matrix 
  * Also, the scale of the parameters has to be addressed 

**Scale Issues for Identification** 

- First, let's consider a simple regression model: 
$$Y_j = B_0 + B_1X_j+\epsilon_J$$
- Recall that the scale of a regression slope ($B_1$) is affected by the scale of both predictor (X) and outcome variable (Y) as follows: 
$$ B_1=r_{XY}(\frac{SD_Y}{SD_X}) $$
- Now, let's consider a one-factor model: 
- Similar to a simple regression model, thescale of a factor loading $\lambda$ is affected by the scale of both predictor ($\xi$) and outcome variable ($x$) as follows:
  * $\lambda = r_{\xi x} (\frac{SD_X}{SD_Y})$

1) If a scale is given to $\xi$, the scale of $\lambda$ will follow 
2) If a scale is given to $\lambda$, the scale of $x$ will follow 
  * Typically, Mplus will assign $\lambda_1 = 1$ to the first indicator 
  
- For a CFA model with 2 or more factors, each factor needs to predict 2 or more observed indicators to fully estimate model parameters 
  * Only applies to simplex models (if it has cross-loadings and correlations, not obvious immediately whether model is identifiable or not)
  
- ESEM for differential item functioning? 
  
**CFA Model with Error Correlations**

- Also, it might be our interest to let some error variances to be correlated 

**Hypothesized 1-factor model** 

- If only 6, then would be enough df but would not be the case by including variances 
- stdyx indicates that both standard deviation and something else are assessed 
`sjlabelled` package has refined version of `read_sav` called `read_spss`

**Mplus output** 

- The number ofparameters include mean-structure parameters 
- It is possible to compute $R^2$

### Higher-order and Bi-factor CFA Models

**Higher-order Factor Models** 
- In some context,we may be interested in developing a model with additional factor 
- Error variances correlated - covariate(s) commonly correlated to the outcome variables not included in the analysis as predictor(s)
- Bifactor model fit indices will be inflated 
  * Likely to see better model fit than we are supposed to 
  * However, doesn't make it useless 
- By giving the asterisk, we are saying that we are overriding the default and instead fix the variance of each latent factor to be 1.0
- Will try to estimate all possible correlations between factors 
- Have to be manually constrained to 0 

### Measurement Invariance 

**Multiple Group CFA**

**Measurement Invariance with Multiple-Group CFA**

- Configural Invariance 
  * Only factor structure is the same 
- Metric Invariance (i.e., weak invariance)
  * Factor loadings are the same 
- Scalar Invariance (i.e., strong invariance)
  * Factor loadings & intercepts are the same 
- Strict Invariance 
  * Factor loadings, intercepts, & residual variances are the same 

- If configural invariance model fits better than weak invariance, we do not have evidence for weak ivnariance 
- configural invariance model will always have lower chisq

- Partial invariance is invariance only for a subset of items 
  * We may be interested in identifying specific items that are different in factor loadings and/or intercepts between groups 
  * This idea is closely realted to DIF
- Latent factor means between groups 
- Latent factor correlations between groups 

- Metric against configural nonsignificant - metric invariance supported, more complex model (configural) is not a better fit

_Configural_
`visual by x1 x2 x3;` factor loadings for the visual factor 
`x1-x9` error variances (not factor loadings!)
`visual; textual; speed` variance between these three latent factors. 
`[x1-x9]` represent intercepts 
`[visual@0 textual@0 speed@0]` represent factor means - fixing to known values, arbitrarily to 0

_Metric_
for metric, factor loadings are assumed to be the same so it is not included in model specification 

_Scalar_
`[x1-x9] (1-9)` indicate items 1-9 same between two groups 

- in the scalar invariance model, where we assume intercept is different between two grouups, if we assume X3 is different between two groups, chisq improves by 19.193, if we assume X7 is different between two groups, chisq improves by 15.022
  * Can conclude that intercepts for items 3 and 7 are not invariant 

_Partial Scalar_
`[x1 x2 x4 x5 x6 x8 x9] (1-7);` 
