---
title: "Class 2 - Mediation Analysis"
date: "`r Sys.Date()`"
output: 
  html_notebook:
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: true
---

```{r message=F, echo = FALSE, warning = FALSE, results='hide'}
# detach all libraries

detachAllPackages <- function() {
  basic.packages <- c("package:stats", "package:graphics", "package:grDevices", "package:utils", "package:datasets", "package:methods", "package:base")
  package.list <- search()[ifelse(unlist(gregexpr("package:", search()))==1, TRUE, FALSE)]
  package.list <- setdiff(package.list, basic.packages)
  if (length(package.list)>0)  for (package in package.list) detach(package,  character.only=TRUE)
}
detachAllPackages()

# function to load libraries

pkgTest <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[,  "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg,  dependencies = TRUE)
  sapply(pkg,  require,  character.only = TRUE)
}

# loading packages 


lapply(c("gsubfn"),pkgTest)
lapply(c("tidyverse"),pkgTest)
lapply(c("sjmisc"),pkgTest)
lapply(c("DT"),pkgTest)
lapply(c("lavaan"),pkgTest)
lapply(c("MplusAutomation"),pkgTest)
lapply(c("psych"),pkgTest)
library(gsubfn)
library(tidyverse)
library(sjmisc)
library(DT)
library(lavaan)
library(MplusAutomation)
library(psych)

if (!require("BiocManager", quietly = TRUE)) install.packages("BiocManager")

BiocManager::install("rhdf5")
```

```{r}
source("mplus.R")
```

- Input not case-sensitive, output has to be in uppercase 
### Mediation Analysis 

- The difference $c-c^\prime$ is considered as the mediation effect, and $c-c^\prime = a \times b$ or $c = a \times b - c^\prime$
- However, the equity holds only for a special case of mediation analysis 
  * Continuous outcome and linear regression 
- Therefore, it is not recommended to evaluate $c - c^\prime$

**Basic Mediation Analysis**

- Data from the discrimination study (Garcia et al., 2010) in Muthen et al. (2016; p.72), as well as Hayes (2013) 
- Subjects (129 females) were presented a story of hoow female attorney lost a promotion to a less qualified male due to discriminatory action 
- Subjects were randomly assigned to two different scenarios: 
  * (1) the female attorney takes no action, and (2) the female attorney protests 
- Variables 
  * Intervention variable (0 = no protest, 1 = protest) (`protest`) predictor 
  * How well the subject likes the female attorney outcome - 7-point likert (`liking`)
  * Perceived appropriateness of the response of the female attorney (`respappr`) mediator 
  * Beliefs in the pervasiveness of sex discrimination in society (`sexism`)
  
```{r}
protest_df <- read.table("protest.txt", sep="")
colnames(protest_df) <- c("sexism", "liking", "respappr", "protest")
psych::describe(protest_df)
```

- By substituting the second equation below to the first equation below, we can get the combined equation:  

Equation 1:
$$ y_i = \beta_0 + \beta_1m_i + \beta_2x_i+\epsilon_{yi}$$
Equation 2: 
$$ m_i = \gamma_0 + \gamma_1x_i+\epsilon_{mi}$$
Combined equation:
$$ y_i = \beta_0 + \beta_1(\gamma_0 + \gamma_1x_i + \epsilon_{mi})+\beta_2x_i+\epsilon_{yi}$$

- Example 1
  * Note that by default, Mplus utilizes a method to estimate a standard error of nonlinearly transformed parameters called the "Delta method". 
  * Delta method analytically approximates the target standard error by assuming the normality of the sampling distribution of the parameter(s). 
    - The indirect effect is a multiplication of two parameters 
  * Mplus syntax for the `model` section: `DV on IV1 IV2` and so on for however many IVs there are 
  
```{r}
sink("ex2.1.inp")
cat('
data:     file = protest.txt; 

variable: names = sexism liking respappr protest; 
          usev = liking respappr protest; 
          
analysis: estimator = mlr; 

model:    liking on respappr protest; 
          respappr on protest; 
          
Model indirect: 
          liking ind protest; 
          
output:   cinterval;
')
sink()

runModels("ex2.1.inp")

modelfit1 <- readModels("ex2.1.out")

modelfit1$parameters$unstandardized
```

**Basic Mediation Analysis with Bootstrap**

```{r}
sink("ex2.1b.inp")
cat('
data:     file = protest.txt; 

variable: names = sexism liking respappr protest; 
          usev = liking respappr protest; 
          
analysis: estimator = ml; 
          bootstrap = 10000;

model:    liking on respappr protest; 
          respappr on protest; 
          
Model indirect: 
          liking ind protest; 
          
output:   cinterval(bootstrap);
')
sink()

runModels("ex2.1b.inp")

modelfit2 <- readModels("ex2.1b.out")

modelfit2$parameters$unstandardized
```

**Mediation Analysis with Covariates** 

- Data from the Maternal Health Project 
- Sample: Mothers who drank at least three drinks a week during the first trimester of their pregnancy + mothers who used alcohol less often ($n=594$)
- Variables 
  * Alcohol use (`momcig3`)
  * Cigarette use (`momalc3`)
  * Child's gender (`male`)
  * Ethnicity (`black`)
  * Head circumference at birth (`hcirc0`)
  * Head circumference at birth 
**Another way to set up mediation analysis** 

```{r}
sink("ex2.1c.inp")
cat('
data:     file = protest.txt; 

variable: names = sexism liking respappr protest; 
          usev = liking respappr protest; 
          
analysis: estimator = ml; 
          bootstrap = 10000;

model:    liking on respappr (b1)
                     protest; 
          respappr on protest (g1); 
          
Model constraint: 
          new(prolik); 
          prolik = g1 * b1;
          
output:   cinterval(bootstrap);
')
sink()

runModels("ex2.1c.inp")

modelfit3 <- readModels("ex2.1c.out")

modelfit3$parameters$unstandardized
```

**Basic Mediation Analysis with** `lavaan`

```{r}
MA.mod01 <- '
liking ~ b1*respappr + protest 
respappr ~ g1*protest 
prolik := b1*g1
'
```

Run the model with bootstrap. 
```{r}
MA.fit01 <- sem(MA.mod01, data=protest_df, se="boot", bootstrap=10000)
```

```{r}
parameterEstimates(MA.fit01, zstat=F, pvalue=F)
```

**Mediation Analysis with Covariates** 

- Data from the Maternal Health Project (Day et al., 1994) is from a sample of mothers who drank at least three drinks a week during the first trimester of their pregnancy + mothers who drank alcohol less often ($n$ = 594). 
- Variables
  * Alcohol use (`momcig3`)
  * Cigarette use (`momalc3`)
  * Child's gender (`male`)
  * Ethnicity (`black`)
  * Head circumference at birth (`hcirc0`)
  * Head circumference at 36 monoths (`hcirc36`)
- The hypothesis is that there are no direct effects of `momalc3` and `momcig3` on `hcirc36`. 

Equation 1: 
$$
y_i = \beta_0+\beta_1 m_i+\beta_2z_{1i}+\beta_3z_{2i}+\epsilon_{yi}
$$
Equation 2: 
$$
m_i=\gamma_0+\gamma_1x_{1i}+\gamma_2x_{2i}+\gamma_3z_{1i}+\gamma_4z_{2i}+\epsilon_{mi}
$$
- Substitute the second equation to the first equation: 
$$
y_i = \beta_0 + \beta_1(\gamma_0 + \gamma_1x_{1i}+\gamma_2x_{2i}+\gamma_3z_{1i}+\gamma_4z_{2i}+\epsilon_{mi}) + \beta_2z_{1i}+\beta_3z_{2i}+\epsilon_{yi}
$$
  * No direct effect of $x$ to $y$ in this case. The indirect effect of $x_1$ to $y$ is $\beta_1 \gamma_1$; the indirect of $x_2$ to $y$ is $\beta_1 \gamma_2$. 
  
  
**Mediation Analysis with multiple Mediators**

- Data from a randomized study of media influence 
- 43 male aned 80 females read a newspaper article and were randomly assigned to one of the two groups 
  * One group was told that the story would appear on the front page of a major newspaper 
  * The other group was told that it would appear in the middle of an economic supplement. 
- Variables
  * Group (0 = less visible; 1 = more visible) (`cond`)
  * Intention to buy sugar and how much (`reaction`)
  * Presumed media influence to others (`pmi`)
  * Importance oof sugar shortage (`import`)
Both direct effect annd mediation effect depends on the value of $z_i$ as a covariate (i.e., moderated by the covariate)

Equation 1: 
$$ 
y_i = \beta_0 + \beta_1m_{1i} + \beta_2m_{2i} + \beta_3{x_i} + \epsilon_{yi}
$$ 
Equation 2: 
$$ 
m_{1i} = \gamma_{10} + \gamma_{11}x_i + \epsilon_{m1i}
m_{2i} = \gamma_{20} + \gamma_{21}x_i + \epsilon_{m2i}

$$
- The direct effect of $x$ to $y$ is $\beta_3$; the indirect effect of $x$ too $y$ through $m_1$ is $\beta_1 \gamma_{11}$, and the indirect effect of $x$ to $y$ through $m_2$ is $\beta_2 \gamma_{21}$.


If the moderation value is lower than .15, mediation effect is not significant as it's within the 95% band around 0. Within .15~+.85 would indicate significance, outside would be not significant. 
- The outvome i

**Moderated Mediation (1)**

```{r}
sink("ex2.23.inp")
cat('
analysis: estimator = ml; 
          bootstrap = 10000; 
model:    remove on agg5 tx agg1 xx; 
          agg5 on tx agg1 xz; 
model indirect: 
  remove MOD agg5 agg1(-1, 1, 0.1) xz tx; 
output: 
  sampstat patterns standardized 
  cinterval(bootstrap);
plot:     type = plot3;
')
sink()

runModels("ex2.23.inp")

modelfit4 <- readModels("ex2.23.out")
```

`model indirect`
- This simple approach works when the effects are moderated only by the moderator variable 
- This approach produces "causally defined" direct and indirect effects
- MOD keyword is added 
- Variables are listed in the following order: 
  * $y$ (outcome)
  * MOD keyword 
  * $m$ (mediator) 
  * $z$ (moderator)
  * $xz$ (interaction)
  * $x$ (exposure variable)
- THe range of $z$ (`agg1`) is provided, because the direct and indirect effects depend on the value of $z$.


`model constraint`
- `LOOP` specifies the range of values for the x-axis of the plot. 
- `PLOT` names the variables to be plotted on the 



**Moderated Mediation (2)** 

```{r}
sink(file = "ex2.25.inp")
cat('
data:     file = teams.txt;
variable: names = dysfunc negtone negexp perform;
          usev = dysfunc negtone negexp perform mz;
define:   mz = negtone*negexp;
analysis: estimator = ml;
          bootstrap = 10000;
model:    perform on negtone dysfunc negexp mz;
          negtone on dysfunc;  
model indirect:
          perform MOD negtone negexp(-0.4,0.6,0.01) mz dysfunc (0.4038 0.035);
output:   cinterval (bootstrap);
plot:     type = plot3;
  ', fill=T)
  sink()
```

Run the model.
```{r}
runModels("ex2.25.inp")
```

Plot the moderated mediation effect by using a function available from `mplus.R`.
```{r}
mplus.plot.moderation("ex2.25.gh5")
```

By using a function available in `mplus.R`, read values of mediation effect, lower and upper bounds of 95% CI at each value of the moderator variable. Note that these values are also reported in the Mplus output file `ex2.25.out`. However, it is nice to see them on R too.
```{r}
tibble(
  negexp = mplus.get.moderation.xvalues("ex2.25.gh5"),
  ests = mplus.get.moderation.estimates("ex2.25.gh5"),
  loci = mplus.get.moderation.lowerci("ex2.25.gh5"),
  upci = mplus.get.moderation.upperci("ex2.25.gh5"))
```

Now, run the same model by using a different approach by using "model constraint" command with "loop" to specify a series of evaluation points. 
```{r}
sink(file = "ex2.25b.inp")
cat('
data:     file = teams.txt;
variable: names = dysfunc negtone negexp perform;
          usev = dysfunc negtone negexp perform mz;
define:   mz = negtone*negexp;
analysis: estimator = ml;
          bootstrap = 10000;
MODEL:    perform on negtone (b1)
          dysfunc (b2)
          negexp (b3)
          mz (b4);
          negtone on dysfunc (g1);
MODEL CONSTRAINT:
          plot(ind);
          loop(negexp, -.4, .6, .01);
          ind = (b1*g1 + b4*g1*negexp) * (.4038 - .035);
output:   cinterval (bootstrap);
plot:     type = plot3;
  ', fill=T)
  sink()
```

Run the model.
```{r}
runModels("ex2.25b.inp")
```

Generate a plot of the moderated mediation effect by using a function available from `mplus.R`. Note that the function is different from the one for the other approach in ex2.25.
```{r}
mplus.plot.loop("ex2.25b.gh5", 1)
```

Similar to the previous example, generate a table of mediation effect, lower and upper bounds of 95% CI at each value of the moderator variable by using a function available in `mplus.R`. Note that this table is not available in the Mplus output file when the "model constraint" approach is used. However, the necessary values are stored in the `.gh5` file, so we can extract the values and summarize as a table on R.

```{r}
tibble(
  negexp = mplus.get.loop.xvalues("ex2.25b.gh5"),
  ests = mplus.get.loop.estimates("ex2.25b.gh5", 1),
  loci = mplus.get.loop.lowerci("ex2.25b.gh5", 1),
  upci = mplus.get.loop.upperci("ex2.25b.gh5", 1)) %>% print(n=101)
```

Read the data.
```{r}
ex2.26.dat <- read.table("xmVx4s1n200rep6.dat")
colnames(ex2.26.dat) <- c("y", "m", "x")
ex2.26.dat <-
  ex2.26.dat %>% mutate(mx = m*x)
psych::describe(ex2.26.dat, fast=T)
```

Set up the Mplus specifications with MplusAutomation. Then, run the model with MplusAutomation.
```{r}
sink(file = "ex2.26.inp")
cat('
data:     file = "xmVx4s1n200rep6.dat";
variable: names = y m x;
          usev = y m x mx;
define:   mx = m*x;
analysis: estimator = ml;
          bootstrap = 10000;
MODEL:    y on m x mx;
          m on x;
MODEL INDIRECT:
          y MOD m mx x (7, 5);
output:   cinterval (bootstrap);
plot:     type = plot3;
  ', fill=T)
  sink()
```

By using a function available in `mplus.R`, read values of mediation effect, lower and upper bounds of 95% CI at each value of the moderator variable. Note that these values are also reported in the Mplus output file `ex2.25.out`. However, it is nice to see them on R too.
```{r}
tibble(
  negexp = mplus.get.moderation.xvalues("ex2.25.gh5"),
  ests = mplus.get.moderation.estimates("ex2.25.gh5"),
  loci = mplus.get.moderation.lowerci("ex2.25.gh5"),
  upci = mplus.get.moderation.upperci("ex2.25.gh5"))
```

Interpretation by Aki: When the moderator negative expressiveness value is negative -.14 or larger, mediation effect is statistically significant. (range of value associated with statistically significant effect)



--- 

**Just Identified Model**

The number of unique elements in the data covariance matrix is $\frac{q(q+1)}{2}$ where $q$ is the order of data covariance matrix. Thus, $df = \frac{q(q+1)}{2}-k$. 

Just identified path analysis model perfectly fits the data with $df=0$. The two example models for the just-identified model below has q=3, so the number of unique elements in the data covariance matrix is 3(3+1)2 = 12/2 = 6. For the model on the left, there are two reg coefficients for X1 predicting Y1 and X2 predicting Y1, covariance between X1 and X2, and error variances for X1, X2, and Y1. Total of 6, so $df=6-6=0$.For the model on the right, there are three reg coefficients for X1 predicting Y1, X1 predicting Y2, and Y1 predicting Y2. Also, there are error variances for X1, Y1, and Y2. Total of 6 again, so $df=6-6=0$. 

**Path Analysis**

Fitness is an exogenous variable as it is both a predictor and outcome variable. Others are endogenous variables. 

**Chi-square Test for One Model** 

- The test evaluates whether the observed covariance matrix is consistent with a specified model. 
- It tests the null hypothesis $H_0: \Sigma =  \widehat{\Sigma}$ where $\Sigma$ is the observed data covariance matrix and $\widehat{\Sigma}$ is the reproduced data covariance matrix. 
  * Note: Just like we 

**Chi-square in Mplus**

Chi-square doesn't make sense for just-identified model with $df=0$ because $\Sigma$ and $\widehat{\Sigma}$ are essentially the same, little to no discrepancy. So although $p = 0.0000$, this does not mean we reject $H_0: \Sigma = \widehat{\Sigma}$. However, it is of use for overidentified model with $df=4$. 

Note: use `$summaries` for extracting model fit information. 

** Chi-square Difference Test for Two Nested Models** 

- Two models are nested if one is a subset of the other. The one on the left-hand side has only one additional parameter than the model on the right-hand side, so the model on the right-hand side is nested in the left-hand side. 
- Chi-square difference test assesses whether the chi-square has been reduced enough to claim that the additional parameter on the left hand side has significantly improved model fit. 

```
pchisq(2.48, 1, lower.tail = F) # not statistically signifcant at alpha of .05
```

- If the test indicates that we do not have evidence for better fit for the more complex model, that suggests that the simpler model is a better (more parsimonious model choice)
- Local fit vs. Global fit 
  * Global fit is the fit of the model, in this case the chisq difference test   * Local fit is the significance of each of the paths in the model 
  
**Modification Index**

Larger modification index means that the additional parameter for the more complex model improved model fit significantly. 

`HARDINES` is predictor, `FITESS` is outcome. Y on X. 

Edit to make to 01.PA3c.inp code
```
OUTPUT: stdyx
        modindices
```


