---
title: "Class 2 - Mediation Analysis"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: cosmo
    toc: yes
    toc_float:
      collapsed: true
---

```{r message=F, echo = FALSE, warning = FALSE, results='hide'}
# detach all libraries

detachAllPackages <- function() {
  basic.packages <- c("package:stats", "package:graphics", "package:grDevices", "package:utils", "package:datasets", "package:methods", "package:base")
  package.list <- search()[ifelse(unlist(gregexpr("package:", search()))==1, TRUE, FALSE)]
  package.list <- setdiff(package.list, basic.packages)
  if (length(package.list)>0)  for (package in package.list) detach(package,  character.only=TRUE)
}
detachAllPackages()

# function to load libraries

pkgTest <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[,  "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg,  dependencies = TRUE)
  sapply(pkg,  require,  character.only = TRUE)
}

# loading packages 


lapply(c("gsubfn"),pkgTest)
lapply(c("tidyverse"),pkgTest)
lapply(c("sjmisc"),pkgTest)
lapply(c("DT"),pkgTest)
lapply(c("lavaan"),pkgTest)
lapply(c("MplusAutomation"),pkgTest)
lapply(c("psych"),pkgTest)
library(gsubfn)
library(tidyverse)
library(sjmisc)
library(DT)
library(lavaan)
library(MplusAutomation)
library(psych)
```

- Input not case-sensitive, output has to be in uppercase 
### Mediation Analysis 

- The difference $c-c^\prime$ is considered as the mediation effect, and $c-c^\prime = a \times b$ or $c = a \times b - c^\prime$
- However, the equity holds only for a special case of mediation analysis 
  * Continuous outcome and linear regression 
- Therefore, it is not recommended to evaluate $c - c^\prime$

**Basic Mediation Analysis**

- Data from the discrimination study (Garcia et al., 2010) in Muthen et al. (2016; p.72), as well as Hayes (2013) 
- Subjects (129 females) were presented a story of hoow female attorney lost a promotion to a less qualified male due to discriminatory action 
- Subjects were randomly assigned to two different scenarios: 
  * (1) the female attorney takes no action, and (2) the female attorney protests 
- Variables 
  * Intervention variable (0 = no protest, 1 = protest) (`protest`) predictor 
  * How well the subject likes the female attorney outcome - 7-point likert (`liking`)
  * Perceived appropriateness of the response of the female attorney (`respappr`) mediator 
  * Beliefs in the pervasiveness of sex discrimination in society (`sexism`)
  
```{r}
protest_df <- read.csv("protest.csv")

psych::describe(protest_df$protest)
```

- By substituting the second equation below to the first equation below, we can get the combined equation:  

Equation 1:
$$ y_i = \beta_0 + \beta_1m_i + \beta_2x_i+\epsilon_{yi}$$
Equation 2: 
$$ m_i = \gamma_0 + \gamma_1x_i+\epsilon_{mi}$$
Combined equation:
$$ y_i = \beta_0 + \beta_1(\gamma_0 + \gamma_1x_i + \epsilon_{mi})+\beta_2x_i+\epsilon_{yi}$$

- Example 1
  * Note that by default, Mplus utilizes a method to estimate a standard error of nonlinearly transformed parameters called the "Delta method". 
  * Delta method analytically approximates the target standard error by assuming the normality of the sampling distribution of the parameter(s). 
    - The indirect effect is a multiplication of two parameters 
  * Mplus syntax for the `model` section: `DV on IV1 IV2` and so on for however many IVs there are 
  
```
sink("ex2.1.inp")
cat('
data:     file = protest.txt; 

variable: names = sexism liking resappr protest; 
          usev = liking respappr protest; 
          
analysis: estimator = mlr; 

model:    liking on resappr protest; 
          resappr on protest; 
          
Model indirect: 
          liking ind protest; 
          
output:   cinterval;
')
sink()
```

```
runModels("protest1.inp")
```

- Total indirect > total as there can be more than 1 indirect effect
- Confidence interval is strictly based on bootstrapping 

- Reject null hypothesis, indirect effect of `respappr` is significant as CI does not include 0

**Mediation Analysis with Covariates** 

- Data from the Maternal Health Project 
- Sample: Mothers who drank at least three drinks a week during the first trimester of their pregnancy + mothers who used alcohol less often ($n=594$)
- Variables 
  * Alcohol use (`momcig3`)
  * Cigarette use (`momalc3`)
  * Child's gender (`male`)
  * Ethnicity (`black`)
  * Head circumference at birth (`hcirc0`)
  * Head circumference at birth 
**Another way to set up mediation analysis** 

**Moderated Mediation (1)**

```
sink("ex2.23.inp")
analysis: estimator = ml; 
          bootstrap = 10000; 
model:    remove on agg5 tx agg1 xx; 
          agg5 on tx agg1 xz; 
model indirect: 
  remove MOD agg5 agg1(-1, 1, 0.1) xz tx; 
output: 
  sampstat patterns standardized 
  cinterval(bootstrap);
plot:     type = plot3;
sink()
```
"model indirect" 
- This simple approach works when 