---
title: "SEM Project 3 Self Evaluation"
author: "Caroline Lee"
date: "`r Sys.Date()`"
output: html_notebook
---

```{r setup, include = FALSE, results='hide'}
# detach all libraries

detachAllPackages <- function() {
  basic.packages <- c("package:stats", "package:graphics", "package:grDevices", "package:utils", "package:datasets", "package:methods", "package:base")
  package.list <- search()[ifelse(unlist(gregexpr("package:", search()))==1, TRUE, FALSE)]
  package.list <- setdiff(package.list, basic.packages)
  if (length(package.list)>0)  for (package in package.list) detach(package,  character.only=TRUE)
}
detachAllPackages()

# function to load libraries

pkgTest <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[,  "Package"])]
  if (length(new.pkg)) 
    install.packages(new.pkg,  dependencies = TRUE)
  sapply(pkg,  require,  character.only = TRUE)
}

# loading packages 

lapply(c("DT"),pkgTest)
lapply(c("haven"),pkgTest)
lapply(c("lavaan"),pkgTest)
lapply(c("MplusAutomation"),pkgTest)
lapply(c("readxl"),pkgTest)
lapply(c("sjlabelled"),pkgTest)
lapply(c("sjmisc"),pkgTest)
lapply(c("texreg"),pkgTest)
lapply(c("tidyverse"),pkgTest)

library(DT)
library(haven)
library(lavaan)
library(MplusAutomation)
library(readxl)
library(sjlabelled)
library(sjmisc)
library(texreg)
library(tidyverse)
```

### <span style="color:orange"> **Revisions for the self evaluation are in orange text.** </span>

### Introduction

Parental Self-Care Scale (PSCS) (Abell, Ryan, & Kamata, 2006) is a 30-item scale that measures ill HIV-positive parents' capacities to manage their own emotional, instrumental, and nursing needs while maintaining family responsibilities. Respondents are asked to rate their belief on their own capacity on each of the 30 items with a 11-point scale (0 = Cannot do at all, 5 =  Probably can do, 10 = Sure I can do). The list of items is in the table at the end of this document.

The data set `PSCS_combined.sav` consists of item responses from 500 HIV-positive parents on items 1-30. Column 1-30 are item responses for the 30 items (`x1-x30`) and Column 31 is the language group (`lang`) variable. There were two groups of parents based on the language form they took (variable `lang`; 1 = Spanish, 2 = English). In this project, you will conduct a series of invariance study analyses for Spanish and English versions of the PSCS scale.  

### Questions
#### Q.1  

(a) Load required packages and read data file in. Use `read_spss()` function in the `sjlabelled` package to read the data file into R. Be sure to include `convert.factors=FALSE` option to prevent the variables to be converted to categorical variables. Show R code.  

```{r}
setwd("C:/Users/48909138/OneDrive - Southern Methodist University/Desktop/SEM-S2024/Project_3")
```

<span style="color:orange"> **Added** </span> `convert.factors = FALSE`
```{r}
PSCS <- read_spss("pscs_combined.sav", convert.factors = FALSE)
```

(b) Generate descriptive statistics for the 30 items using `descr()` function in `sjmisc` package. Confirm that there is no irregularity in the data, such as out-of-range scores and excessive missing data.  

```{r}
descr(PSCS[,1:30])
```

(c) Generate a frequency table for the [lang] variable using `frq()` function in `sjmisc` package. Confirm that there is no irregularity in the data, such as out-of-range scores and excessive missing data.  

<span style="color:orange"> **Fixed to only include `lang` as opposed to excluding `lang`** </span>  

```{r}
PSCS %>% select(lang) %>% frq()
```

<span style="color:orange"> **There was no irregularity in the data** </span>  

(d) Using `prepareMplusData()` function in the `MplusAutomation` package, generate an external data file for Mplus use.

```{r results='hide'}
PSCS_dat <- prepareMplusData(PSCS, "PSCS.dat")
```

#### Q.2
Using Mplus, fit a single-group 3-factor CFA model with ULI scaling strategy. Show all R code.   

(a) Generate an external Mplus input file for a 3-factor CFA model with all 30 items.  

```{r}
sink(file = "PSCS.inp")
cat('
DATA:     file = "PSCS.dat";
VARIABLE: names = x1 x2 x3 x4 x5 x6 x7 x8 x9 x10
                  x11 x12 x13 x14 x15 x16 x17 x18 x19 x20 
                  x21 x22 x23 x24 x25 x26 x27 x28 x29 x30 lang;
          usev = x1 x2 x3 x4 x5 x6 x7 x8 x9 x10
                 x11 x12 x13 x14 x15 x16 x17 x18 x19 x20 
                 x21 x22 x23 x24 x25 x26 x27 x28 x29 x30;
          missing = .;
MODEL:    emotion by x1 x2 x3 x4 x5 x6 x7 x8 x9 x10;
          instrum by x11 x12 x13 x14 x15 x16 x17 x18 x19 x20;
          nursing by x21 x22 x23 x24 x25 x26 x27 x28 x29 x30;
OUTPUT:   stdyx modindices;
  ', fill=T)
sink()
```

(b) Run the input file by the `runModels()` function in the `MplusAutomation` package.  

```{r}
runModels("PSCS.inp")
```

(c) By using `readModels()` function in the `MplusAutomation`, read the output file. Then, extract CFI, TLI, RMSEA, and 90% CI of RMSEA. Comment on the model fit based on these values.  

```{r warning=FALSE}
PSCS_fit <- readModels("pscs.out")
```

<span style="color:orange"> **Changed from** </span> `PSCS_fit$summaries %>% unlist()` <span style="color:orange"> **(which provided much more than just CFI, TL, RMSEA, and 90% CI of RMSEA)** </span> 
```{r}
PSCS_fit$summaries %>% as_tibble() %>% select(CFI, TLI, RMSEA_Estimate, RMSEA_90CI_LB, RMSEA_90CI_UB)
```
- We can conclude that the model fit is good based on the following paramters. 
  * CFI is 0.987 and >.95, indicating good fit.
  * TLI is 0.986 and >.95, indicating good fit.
  * RMSEA is 0.021 and $\leq$.06, indicating good fit.
  * Lower bound of 90% CI for RMSEA is 0.014 and $\leq$.05, indicating good fit.
  * Upper bound of 90% CI for RMSEA is 0.027 and <.07, indicating good fit. 

(d) Extract standardized factor loadings.   

```{r}
PSCS_fit$parameters$stdyx.standardized
```

#### Q.3
Now, use Mplus to conduct one Mplus run of overall invariance tests between Spanish and English groups. Show all R code. 

(a) Show R code to generate an external Mplus input file for the analysis. Use ULI scaling strategy. 

```{r}
PSCS %>% 
  mutate(PSCS_MI=recode(lang, 
                        "Spanish" = 1,
                        "English" = 2)) %>% 
  prepareMplusData("PSCS_MI.dat")
```

```{r}
sink('PSCS_MI.inp')
cat('
DATA:      file = "PSCS_MI.dat";
VARIABLE:  names = x1 x2 x3 x4 x5 x6 x7 x8 x9 x10
                   x11 x12 x13 x14 x15 x16 x17 x18 x19 x20 
                   x21 x22 x23 x24 x25 x26 x27 x28 x29 x30 lang;
           grouping is lang (1=Spanish 2=English);
ANALYSIS:  model = configural metric scalar;
MODEL:     emotion by x1 x2 x3 x4 x5 x6 x7 x8 x9 x10;
           instrum by x11 x12 x13 x14 x15 x16 x17 x18 x19 x20;
           nursing by x21 x22 x23 x24 x25 x26 x27 x28 x29 x30;
', fill=T)
sink()
```

```{r}
runModels("PSCS_MI.inp")
```

(b) By using `readModels()` function in the `MplusAutomation`, read the output file. Then extract the Mplus results that show model comparisons between configural, metric, and scalar invariance models. 

```{r, warning=FALSE}
PSCS.MI.fit <- readModels("pscs_mi.out")
```

```{r}
PSCS.MI.fit$invariance_testing
```

(c) Briefly define what metric (weak) invariance means. Then, state whether weak invariance was supported by the result. State why or why not. Evaluate both Chi-square difference test and CFI change.  

- Metric invariance means that the factor loadings and factor structure of PCSC data is the same across the two groups (i.e., Spanish vs. English language groups).
- The configural model (CFI = 0.971, $\chi^2$ = 1007.503) did not improve the model fit compared to the metric model (the more restrictive model; CFI = 0.969, $\chi^2$ = 1042.748). Therefore,  there is evidence for the metric invariance. 

(d) Briefly define what scalar (strong) invariance means. Then, state whether strong invariance was supported by the result. State why or why not. Evaluate both Chi-square difference test and CFI change.  

- Scalar invariance means that the factor loadings, intercepts, and structure are same across the two groups (i.e., Spanish vs. English language groups) for the PSCS dataset. 
- The metric model (CFI = 0.969; $\chi^2$ = 1042.748) did improve the model fit compared to the scalar model (the more restrictive model; CFI = 0.962, $\chi^2$ = 1093.370). Therefore, there is no evidence for the scalar invariance. 

#### Q.4
Now, fit various invariance models with multiple group analysis by Mplus. The goal here is to match the results for all models from the one-step approach in Q.2. For each of the model below, attach the R code to generate the Mplus syntax. Then, run the model by using the `runModels()` function, and read the results by using the `readModels()` function. Then, extract  df, chi-squre statistic, CFI, TLI, AIC, BIC, and aBIC. Verify these values match the results from Q.2.   

(a) Configural invariance model.  

```{r}
sink('PSCS_MI_Configural.inp')
cat('
DATA:       file = "PSCS_MI.dat";
VARIABLE:   names = x1 x2 x3 x4 x5 x6 x7 x8 x9 x10
                   x11 x12 x13 x14 x15 x16 x17 x18 x19 x20 
                   x21 x22 x23 x24 x25 x26 x27 x28 x29 x30 lang;
            grouping is lang (1=Spanish 2=English);
MODEL:      emotion by x1 x2 x3 x4 x5 x6 x7 x8 x9 x10;
            instrum by x11 x12 x13 x14 x15 x16 x17 x18 x19 x20;
            nursing by x21 x22 x23 x24 x25 x26 x27 x28 x29 x30;
            x1-x30;
            emotion; instrum; nursing;
            [x1-x30];
            [emotion@0 instrum@0 nursing@0];
MODEL English:    
            emotion by x1@1 x2 x3 x4 x5 x6 x7 x8 x9 x10;
            instrum by x11@1 x12 x13 x14 x15 x16 x17 x18 x19 x20;
            nursing by x21@1 x22 x23 x24 x25 x26 x27 x28 x29 x30;
            x1-x30;
            emotion; instrum; nursing;
            [x1-x30];
            [emotion@0 instrum@0 nursing@0];
', fill=T)
sink()
```

```{r}
runModels("PSCS_MI_Configural.inp")
```

```{r, warning=FALSE}
PSCS.MI.Configural.fit <- readModels("pscs_mi_configural.out")
```

```{r}
PSCS.MI.Configural.fit$summaries %>% unlist()
```

(b) Metric (weak) invariance model.  

```{r}
sink('PSCS_MI_Metric.inp')
cat('
DATA:       file = "PSCS_MI.dat";
VARIABLE:   names = x1 x2 x3 x4 x5 x6 x7 x8 x9 x10
                   x11 x12 x13 x14 x15 x16 x17 x18 x19 x20 
                   x21 x22 x23 x24 x25 x26 x27 x28 x29 x30 lang;
            grouping is lang (1=Spanish 2=English);
MODEL:      emotion by x1 x2 x3 x4 x5 x6 x7 x8 x9 x10;
            instrum by x11 x12 x13 x14 x15 x16 x17 x18 x19 x20;
            nursing by x21 x22 x23 x24 x25 x26 x27 x28 x29 x30;
            x1-x30;
            emotion; instrum; nursing;
            [x1-x30];
            [emotion@0 instrum@0 nursing@0];
MODEL English:    
            x1-x30;
            emotion; instrum; nursing;
            [x1-x30];
            [emotion@0 instrum@0 nursing@0];
', fill=T)
sink()
```

```{r}
runModels("PSCS_MI_Metric.inp")
```

```{r, warning=FALSE}
PSCS.MI.Metric.fit <- readModels("pscs_mi_metric.out")
```

```{r}
PSCS.MI.Metric.fit$summaries %>% unlist()
```
 
(c) Scalar (strong) invariance model.  

```{r}
sink('PSCS_MI_Scalar.inp')
cat('
DATA:       file = "PSCS_MI.dat";
VARIABLE:   names = x1 x2 x3 x4 x5 x6 x7 x8 x9 x10
                   x11 x12 x13 x14 x15 x16 x17 x18 x19 x20 
                   x21 x22 x23 x24 x25 x26 x27 x28 x29 x30 lang;
            grouping is lang (1=Spanish 2=English);
MODEL:      emotion by x1 x2 x3 x4 x5 x6 x7 x8 x9 x10;
            instrum by x11 x12 x13 x14 x15 x16 x17 x18 x19 x20;
            nursing by x21 x22 x23 x24 x25 x26 x27 x28 x29 x30;
            x1-x30;
            emotion; instrum; nursing;
            [x1-x30] (1-30);
            [emotion instrum nursing];
MODEL English:    
            x1-x30;
            [x1-x30] (1-30);
            [emotion instrum nursing];
', fill=T)
sink()
```

```{r}
runModels("PSCS_MI_Scalar.inp")
```

```{r, warning=FALSE}
PSCS.MI.Scalar.fit <- readModels("pscs_mi_scalar.out")
```

```{r}
PSCS.MI.Scalar.fit$summaries %>% unlist()
```
```{r}
rbind(PSCS.MI.Configural.fit$summaries %>% as_tibble() %>% 
        select(ChiSqM_Value, ChiSqM_DF, CFI, TLI, AIC, BIC, aBIC), 
      PSCS.MI.Metric.fit$summaries %>% as_tibble() %>% 
        select(ChiSqM_Value, ChiSqM_DF, CFI, TLI, AIC, BIC, aBIC),
      PSCS.MI.Scalar.fit$summaries %>% as_tibble() %>% 
        select(ChiSqM_Value, ChiSqM_DF, CFI, TLI, AIC, BIC, aBIC)) %>% 
  mutate(ChiSqDif = lead(ChiSqM_Value) - ChiSqM_Value,
         DfDif = lead(ChiSqM_DF) - ChiSqM_DF,
         pvalue = (1-pchisq(ChiSqDif, DfDif)) %>% round(3),
         Model = c("Configural" ,"Metric", "Scalar")) %>% 
  relocate(Model)
```

#### Q.5
It is suspected that the fit of the scalar (strong) invariance model would improve by varying their intercepts between groups for items 2 and 5. Therefore, we will test partial-strong invariance models targeting the intercepts of the two items in this question. Show all R code.  

(a) Specify a partial scalar-invariance model, where intercepts for both items 2 & 5 are not invariant between the two groups. Show R code to generate Mplus syntax file.  

```{r}
sink('PSCS_MI_Partial_Scalar.inp')
cat('
DATA:       file = "PSCS_MI.dat";
VARIABLE:   names = x1 x2 x3 x4 x5 x6 x7 x8 x9 x10
                   x11 x12 x13 x14 x15 x16 x17 x18 x19 x20 
                   x21 x22 x23 x24 x25 x26 x27 x28 x29 x30 lang;
            grouping is lang (1=Spanish 2=English);
MODEL:      emotion by x1 x2 x3 x4 x5 x6 x7 x8 x9 x10;
            instrum by x11 x12 x13 x14 x15 x16 x17 x18 x19 x20;
            nursing by x21 x22 x23 x24 x25 x26 x27 x28 x29 x30;
            x1-x30;
            emotion; instrum; nursing;
            [x1-x30] (1-30);
            [x2 x5];
            [emotion instrum nursing];
MODEL English:    
            x1-x30;
            [x1-x30] (1-30);
            [x2 x5]
            [emotion instrum nursing];
', fill=T)
sink()
```

```{r}
runModels("PSCS_MI_Partial_Scalar.inp")
```

(b) Read the results by using `readModels()` function. Then, extract df, chi-squre statistic, CFI, TLI, AIC, BIC, and aBIC to compare the results to the weak invariance model. In the table, include the model labels, as well as the Chi-square difference test to test whether this partial scalar-invariance model is supported.  

```{r, warning=FALSE}
PSCS.MI.PartialScalar.fit <- readModels("pscs_mi_partial_scalar.out")
```

```{r}
PSCS.MI.PartialScalar.fit$summaries %>% unlist()
```

```{r}
rbind(PSCS.MI.Metric.fit$summaries %>% as_tibble() %>% 
        select(ChiSqM_Value, ChiSqM_DF, CFI, TLI, AIC, BIC, aBIC), 
      PSCS.MI.PartialScalar.fit$summaries %>% as_tibble() %>% 
        select(ChiSqM_Value, ChiSqM_DF, CFI, TLI, AIC, BIC, aBIC)) %>%
  mutate(ChiSqDif = lead(ChiSqM_Value) - ChiSqM_Value,
         DfDif = lead(ChiSqM_DF) - ChiSqM_DF,
         pvalue = (1-pchisq(ChiSqDif, DfDif)) %>% round(3),
         Model = c("Metric", "Partial Scalar")) %>% 
  relocate(Model)
```

(c) Conduct the Chi-square difference test by manually entering the values to the `pchisq()` function to verify the p-value in the table. 

```{r}
1 - pchisq((1072.540-1042.748), 25)
```

(d) Comment whether this partial-scalar invariance model is supported by Chi-square difference test, information criteria and other fit indices.  

- As the metric model does not significantly improve the model fit and as the information criterion indicates the partial scalar model is preferred, there is evidence for partial scalar invariance except for items 2 and 5.

(e) Repeat (a)-(d) for another partial-scalar invariance model with only item 2 intercept is variant. 

```{r}
sink('PSCS_MI_Partial_Scalar_02.inp')
cat('
DATA:       file = "PSCS_MI.dat";
VARIABLE:   names = x1 x2 x3 x4 x5 x6 x7 x8 x9 x10
                   x11 x12 x13 x14 x15 x16 x17 x18 x19 x20 
                   x21 x22 x23 x24 x25 x26 x27 x28 x29 x30 lang;
            grouping is lang (1=Spanish 2=English);
MODEL:      emotion by x1 x2 x3 x4 x5 x6 x7 x8 x9 x10;
            instrum by x11 x12 x13 x14 x15 x16 x17 x18 x19 x20;
            nursing by x21 x22 x23 x24 x25 x26 x27 x28 x29 x30;
            x1-x30;
            emotion; instrum; nursing;
            [x1-x30] (1-30);
            [x2];
            [emotion instrum nursing];
MODEL English:    
            x1-x30;
            [x1-x30] (1-30);
            [x2]
            [emotion instrum nursing];
', fill=T)
sink()
```

```{r}
runModels("PSCS_MI_Partial_Scalar_02.inp")
```
 
```{r, warning=FALSE}
PSCS.MI.PartialScalar02.fit <- readModels("pscs_mi_partial_scalar_02.out")
```

```{r}
PSCS.MI.PartialScalar02.fit$summaries %>% unlist()
```

```{r}
rbind(PSCS.MI.Metric.fit$summaries %>% as_tibble() %>% 
        select(ChiSqM_Value, ChiSqM_DF, CFI, TLI, AIC, BIC, aBIC), 
      PSCS.MI.PartialScalar02.fit$summaries %>% as_tibble() %>% 
        select(ChiSqM_Value, ChiSqM_DF, CFI, TLI, AIC, BIC, aBIC)) %>%
  mutate(ChiSqDif = lead(ChiSqM_Value) - ChiSqM_Value,
         DfDif = lead(ChiSqM_DF) - ChiSqM_DF,
         pvalue = (1-pchisq(ChiSqDif, DfDif)) %>% round(3),
         Model = c("Metric", "Partial Scalar")) %>% 
  relocate(Model)
```
```{r}
1 - pchisq((1082.897 - 1042.748), 26)
```
- As the metric model significantly improves the model fit and as the information criterion indicates the metric model is preferred, there is no evidence for partial scalar invariance except for items 2.

(f) Repeat (a)-(d) for another partial-scalar invariance model with only item 5 intercept is variant.

```{r}
sink('PSCS_MI_Partial_Scalar_05.inp')
cat('
DATA:       file = "PSCS_MI.dat";
VARIABLE:   names = x1 x2 x3 x4 x5 x6 x7 x8 x9 x10
                   x11 x12 x13 x14 x15 x16 x17 x18 x19 x20 
                   x21 x22 x23 x24 x25 x26 x27 x28 x29 x30 lang;
            grouping is lang (1=Spanish 2=English);
MODEL:      emotion by x1 x2 x3 x4 x5 x6 x7 x8 x9 x10;
            instrum by x11 x12 x13 x14 x15 x16 x17 x18 x19 x20;
            nursing by x21 x22 x23 x24 x25 x26 x27 x28 x29 x30;
            x1-x30;
            emotion; instrum; nursing;
            [x1-x30] (1-30);
            [x5];
            [emotion instrum nursing];
MODEL English:    
            x1-x30;
            [x1-x30] (1-30);
            [x5]
            [emotion instrum nursing];
', fill=T)
sink()
```

```{r}
runModels("PSCS_MI_Partial_Scalar_05.inp")
```
 
```{r, warning=FALSE}
PSCS.MI.PartialScalar05.fit <- readModels("pscs_mi_partial_scalar_05.out")
```

```{r}
PSCS.MI.PartialScalar05.fit$summaries %>% unlist()
```

```{r}
rbind(PSCS.MI.Metric.fit$summaries %>% as_tibble() %>% 
        select(ChiSqM_Value, ChiSqM_DF, CFI, TLI, AIC, BIC, aBIC), 
      PSCS.MI.PartialScalar05.fit$summaries %>% as_tibble() %>% 
        select(ChiSqM_Value, ChiSqM_DF, CFI, TLI, AIC, BIC, aBIC)) %>%
  mutate(ChiSqDif = lead(ChiSqM_Value) - ChiSqM_Value,
         DfDif = lead(ChiSqM_DF) - ChiSqM_DF,
         pvalue = (1-pchisq(ChiSqDif, DfDif)) %>% round(3),
         Model = c("Metric", "Partial Scalar")) %>% 
  relocate(Model)
```
```{r}
1 - pchisq((1079.892 - 1042.748), 26)
```
- As the metric model does not significantly improve the model fit and as the information criterion indicates the partial scalar model is preferred, there is evidence for partial scalar invariance except for item 5.


