---
title: "Use R Notebook to run MplusAutomation and lavaan"
output: html_notebook
---

### Preparation
Load Required Package.
```{r}
library(tidyverse)
library(MplusAutomation)
library(lavaan)
```

### Our First `MplusAutomation` Attempt
Let's try to run a regression, where [y] is the outcome and [x1] and [x2] are regressors.  

#### Step 1: 
Generate an Mplus syntax file.
```{r}
sink("RegByMA00.inp")
cat('
DATA:       FILE IS ex3.9.dat;
VARIABLE:   NAMES ARE y x1 x2;
MODEL:      y on x1 x2;
OUTPUT:     STDYX;
')
sink()
```

#### Step 2: 
Use `runModels()` function in the `MplusAutomation` package to run the specified syntax file.
```{r}
runModels("RegByMA00.inp")
```
We can take a look at the Mplus output file `regbyma00.out` to make sure that it ran without any issue.  

#### Step 3: 
Use `readModels()` function in the `MplusAutomation` package to read the specified output file.
```{r}
fitMA00 <- readModels("regbyma00.out")
```

This is it!  

Let's take a look at the structure of `fitMA00`.

#### Step 4:
Various ways to see the key output on R.
```{r}
fitMA00$parameters$unstandardized
```

```{r}
fitMA00$parameters$stdyx.standardized
```
Use `screenreg()` function in `texreg` package.
```{r}
texreg::screenreg(fitMA00, 
                  summaries = c("Observations", "CFI", "SRMR"), 
                  single.row=TRUE)
```

## Running lavaan package

#### Step 0:
Import data.
```{r}
dat02 <- read.table("ex3.9.dat")
colnames(dat02) <- c("y", "x1", "x2")
```

#### Step 1:
Generate a model object. This is created within R.
```{r}
mod01 <- 'y ~ 1 + x1 + x2'
```

#### Step 2:
Running the model
```{r}
fit01 <- sem(mod01, data=dat02)
```

#### Step 3:
Various ways to extract the results
```{r}
summary(fit01)
```

```{r}
parameterEstimates(fit01)
```

```{r}
standardizedSolution(fit01)
```

